---
title: "p8105_hw3_cy2588"
author: Chufeng Yang
date: 10/7/2020
output: github_document
---

I'm an R Markdown document! 

## Problem 1

```{r set up}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
data("instacart")
data("ny_noaa")

knitr::opts_chunk$set(
  fig.height = 8,
  fig.width = 8,
  fig.asp = .6,
  out.width = "90%"
)
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

* How many aisles are there, and which aisles are the most items ordered from?  

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```
There are 134 aisles, and fresh vegetables are the most items ordered from.


* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers.  

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	) %>% 
knitr::kable()
```

## Problem 2

Load, tidy, and otherwise wrangle the data.
```{r warning=FALSE}
accel_df = 
  read.csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity.",
    values_to = "activity_count"
    ) %>% 
  mutate(
    weekday_vs_weekend = case_when(
      day %in% c("Saturday","Sunday") ~ "weekend",
      !(day %in% c("Saturday","Sunday")) ~ "weekday"
      )
    ) %>% 
  mutate(
    weekday_vs_weekend = as.factor(weekday_vs_weekend),
    day_id = as.factor(day_id), 
    week = as.factor(week),
    day = as.factor(day),
    minute = as.numeric(minute),
    activity_count = as.numeric(activity_count)
    )%>%
 mutate(day = forcats::fct_relevel(day,
                                   c("Monday","Tuesday","Wednesday","Thursday","Friday", "Saturday","Sunday"))
         ) %>% 
  relocate(day_id, week, weekday_vs_weekend) %>% 
  group_by(week) %>% 
  arrange(day, .by_group = TRUE)
```
For now, in the final dataset, there are variables include `r names(accel_df)`, with a total of `r nrow(accel_df)` rows and `r ncol(accel_df)` columns, there are `r nrow(accel_df)` observations.  

Traditional analyses of accelerometer data focus on the total activity over the day.Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
accel_df %>% 
  group_by(week,day) %>% 
  summarise(total_activity = sum(activity_count))%>% 
  pivot_wider(
    names_from = "day", 
    values_from = "total_activity")%>% 
knitr::kable()
```
It seems like activities from Tuesday to Friday are basically stable during these 5 weeks, however, the activities are significant lower on Saturday and Sunday in week 4 and 5, especially Saturday.


Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.
```{r}
accel_df %>% 
  group_by(day_id, day, week) %>% 
  ggplot(aes(x = minute, 
             y = activity_count, 
             color = day, 
             group = day_id)) + 
  geom_point(alpha = .05)+
  stat_smooth(se = F, 
              method = "loess") + 
  labs(title = "Activities plot",
       caption = "Data from accelerometer data collected on a 63 year-old male with BMI 25")+
  scale_x_continuous(name ="Hour of day" ,
                     breaks = c(seq(120,1440,120)), 
                     labels = c("2h", "4h", "6h", "8h", "10h", "12h","14h", "16h", "18h","20h", "22h", "24h")) +
  scale_y_continuous(name = "activity count", 
                      trans = "log10")
```
Based on this graph, it's seems that daily active time of this 63 year-old male is basically from around 8am to 10pm, and for weekday, his activity counts would be higher and with longer active time.


## Problem 3

### 1
Do some data cleaning.
```{r}
ny_noaa_df = 
  ny_noaa %>%
  separate(date, into = c("year","month","day")) %>%
  mutate_at(vars(year, month, day), as.factor) %>%
  mutate_at(vars(prcp, snow, tmax, tmin), as.numeric) %>%
  mutate(
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10,
  )
skimr::skim_without_charts(ny_noaa_df)
```

For snowfall, what are the most commonly observed values? Why?
```{r}
ny_noaa_df %>%
  filter(snow != "NA") %>%
  count(snow) %>%
  filter(snow >= "0")%>%
  arrange(desc(n))
```
The most commonly observed value is 0, because It didn't snow in most of days.

### 2
Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?
```{r}
ny_noaa_df %>% 
  filter(month %in% c("01","07")) %>% 
  group_by(year, month, id) %>% 
  summarize(
    mean_tmax = mean(tmax),na.rm = T
  ) %>%
  drop_na() %>%
  ggplot(aes(x = year, 
             y = mean_tmax,
             color = id,
             group = id)) + 
  geom_point(alpha = 0.3) + 
  geom_path(alpha = 0.3) +
  facet_grid(~ month) +
  theme(legend.position = 'none',
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(title = "Average max temperature in January and in July",
       caption = "Data from NOAA National Climatic Data Center",
       x = "Year",
       y = "Temperature(C)")
```
There's a increasing of tmax in January from 1981 to 1993, but not continue increase after that. The tmax of July are relatively stable. And all the temperature data basically follow the same variation, and relatively keep same differences between stations.
There do have some outliers, such as temperature of January in 1982 or temperature of July in 1988 and 2004. They may caused by abnormal changes in the weather or just wrong data.
Even if things are messy, are some stations usually higher or lower than others across years? Are some years higher or lower than you might expect? 

### 3
plot 1: tmax vs tmin for the full dataset.
```{r}
temp_plot = 
  ny_noaa_df %>%
   ggplot(aes(x = tmax, y = tmin)) + 
   geom_bin2d()+
   labs(title = "tmax vs tmin",
       x = "max temperature(C)",
       y = "min temperature(C)")
```



plot 2: the distribution of snowfall values greater than 0 and less than 100 separately by year.
```{r fig.height=12,fig.width=8}
snow_plot = 
  ny_noaa_df %>%
  filter(between(snow,1,100))%>%
  ggplot(aes(x= year, y = snow))+
  geom_violin()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))+
   labs(title = "distribution of snowfall separately by year",
       caption = "Data from NOAA National Climatic Data Center",
       x = "Year",
       y = "Snowfall(mm)")
```

combine panels
```{r warning=F}
temp_plot / snow_plot
```





